---
title: "R Notebook"
output: html_notebook
---

* Convert original input to numbers
* Get patch distribution for each color
* Count colors and re-balance to reduce waste
* Re-generate input until one works, if none work, go back to step 3

```{r}
df <- read.csv("/Users/guslipkin/Documents/AoC-Home-Edition/Quilting/input.csv", colClasses = rep("character", 10), header = FALSE)
masterDF <- df
```

Convert original input to numbers while also making a list of all patch sizes for each color
```{r, warning=FALSE}
numberDF <- df

redSize <- c()
orangeSize <- c()
yellowSize <- c()
greenSize <- c()
blueSize <- c()
indigoSize <- c()
violetSize <- c()
whiteSize <- c()

colors <- c("r", "o", "y", "g", "b", "i", "v")
colorCount <- rep(0, 7)

boxSize <- 0

oldColor <- "z"

for(r in 2:(nrow(df) - 1)) {
  for(c in 2:(ncol(df) - 1)) {
    currentColor <- df[r, c]
    
    if(!is.na(currentColor)) {
      for(place in c:(ncol(df) - 1)) {
        if(!is.na(df[r, place]) & currentColor == df[r, place])
          boxSize <- boxSize + 1
        else
          break
      }
      
      testColor <- as.data.frame(df[r:(r + boxSize), c:(c + boxSize)])[1, 1]
      if (testColor == "r")
        redSize <- append(redSize, boxSize)
      if (testColor == "o")
        orangeSize <- append(orangeSize, boxSize)
      if (testColor == "y")
        yellowSize <- append(yellowSize, boxSize)
      if (testColor == "g")
        greenSize <- append(greenSize, boxSize)
      if (testColor == "b")
        blueSize <- append(blueSize, boxSize)
      if (testColor == "i")
        indigoSize <- append(indigoSize, boxSize)
      if (testColor == "v")
        violetSize <- append(violetSize, boxSize)
      if (testColor == "w")
        whiteSize <- append(whiteSize, boxSize)
      
      boxSize <- boxSize - 1
      
      
      colorCount[match(df[r, c], colors)] <- colorCount[match(df[r, c], colors)] + 1
      numberDF[r:(r + boxSize), c:(c + boxSize)] <- (boxSize + 1)
      df[r:(r + boxSize), c:(c + boxSize)] <- NA
      boxSize <- 0
      oldColor <- currentColor
    }
    
  }
}

df <- masterDF
colorList <- list(redSize, orangeSize, yellowSize, greenSize, blueSize, indigoSize, violetSize, whiteSize)

newList <- list()
for(i in 1:length(colorList))
  newList <- append(newList, list(sort(as.vector(colorList[[i]][]), decreasing = TRUE)))
colorList <- newList
```

Create the bolt of fabric and get the last row
```{r, warning=FALSE}
width <- floor((ncol(df) - 2) * .625)
redDF <- data.frame(matrix(NA, nrow=sum(redSize), ncol=width))
orangeDF <- data.frame(matrix(NA, nrow=sum(orangeSize), ncol=width))
yellowDF <- data.frame(matrix(NA, nrow=sum(yellowSize), ncol=width))
greenDF <- data.frame(matrix(NA, nrow=sum(greenSize), ncol=width))
blueDF <- data.frame(matrix(NA, nrow=sum(blueSize), ncol=width))
indigoDF <- data.frame(matrix(NA, nrow=sum(indigoSize), ncol=width))
violetDF <- data.frame(matrix(NA, nrow=sum(violetSize), ncol=width))

part2 <- function(xdf, xcolorList) {
  for (i in 1:length(xcolorList)) {
    breakFlag <- FALSE
    boxSize <- as.numeric(unlist(xcolorList)[i]) - 1
    
    #iterate over cells one by one, fill block as they come, reverse of part 1, kinda
    #for(r in 1:9) {
    for (r in 1:nrow(xdf)) {
      for (c in 1:ncol(xdf)) {
        if (r + boxSize <= nrow(xdf) & c + boxSize <= ncol(xdf)) {
          if (!(FALSE %in% is.na(xdf[r:(r + boxSize), c:(c + boxSize)]))) {
            xdf[r:(r + boxSize), c:(c + boxSize)] <- (boxSize + 1)
            breakFlag <- TRUE
            break
          }
        }
        
      }
      if (breakFlag)
        break
    }
    
  }
  
  return(xdf)
}

redDF <- part2(redDF, colorList[[1]][])
orangeDF <- part2(orangeDF, colorList[[2]][])
yellowDF <- part2(yellowDF, colorList[[3]][])
greenDF <- part2(greenDF, colorList[[4]][])
blueDF <- part2(blueDF, colorList[[5]][])
indigoDF <- part2(indigoDF, colorList[[6]][])
violetDF <- part2(violetDF, colorList[[7]][])

redDF <- redDF[rowSums(is.na(redDF)) != ncol(redDF),]
orangeDF <- orangeDF[rowSums(is.na(orangeDF)) != ncol(orangeDF),]
yellowDF <- yellowDF[rowSums(is.na(yellowDF)) != ncol(yellowDF),]
greenDF <- greenDF[rowSums(is.na(greenDF)) != ncol(greenDF),]
blueDF <- blueDF[rowSums(is.na(blueDF)) != ncol(blueDF),]
indigoDF <- indigoDF[rowSums(is.na(indigoDF)) != ncol(indigoDF),]
violetDF <- violetDF[rowSums(is.na(violetDF)) != ncol(violetDF),]

lastRed <- redDF[rowSums(is.na(redDF)) > 0,]
lastOrange <- orangeDF[rowSums(is.na(orangeDF)) > 0,]
lastYellow <- yellowDF[rowSums(is.na(yellowDF)) > 0,]
lastGreen <- greenDF[rowSums(is.na(greenDF)) > 0,]
lastBlue <- blueDF[rowSums(is.na(blueDF)) > 0,]
lastIndigo <- indigoDF[rowSums(is.na(indigoDF)) > 0,]
lastViolet <- violetDF[rowSums(is.na(violetDF)) > 0,]
```

```{r, eval=FALSE}
write.table(redDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",")
write.table(orangeDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
write.table(yellowDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
write.table(greenDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
write.table(blueDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
write.table(indigoDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
write.table(violetDF, "../Quilting/colors.csv", col.names = FALSE, row.names = FALSE, sep = ",", append = TRUE)
```

Create a vector of the color distributions
```{r, warning=FALSE}
paste("width:", width)
red <- as.numeric(as.vector(lastRed[nrow(lastRed),]))
orange <- as.numeric(as.vector(lastOrange[nrow(lastOrange),]))
yellow <- as.numeric(as.vector(lastYellow[nrow(lastYellow),])) 
green <- as.numeric(as.vector(lastGreen[nrow(lastGreen),]))
blue <- as.numeric(as.vector(lastBlue[nrow(lastBlue),]))
indigo <- as.numeric(as.vector(lastIndigo[nrow(lastIndigo),])) 
violet <- as.numeric(as.vector(lastViolet[nrow(lastViolet),]))
white <- whiteSize

leftover <- data.frame(red, orange, yellow, green, blue, indigo, violet)
leftover

spaces <- colSums(is.na(leftover))
```

Redistribute the colors
```{r}
for(i in 1:length(spaces)) {
  #find the min
  min <- width
  minPlace <- 0
  for (j in 1:length(spaces))
    if (spaces[[j]] < min & spaces[[j]] != 0) {
      min <- spaces[[j]]
      minPlace <- j
    }
  
  #find the max
  max <- 0
  maxPlace <- 0
  for (j in 1:length(spaces))
    if (spaces[[j]] > max & spaces[[j]] != width) {
      max <- spaces[[j]]
      maxPlace <- j
    }
  
  if (maxPlace == 0 | minPlace == 0 | maxPlace == minPlace)
    break
  
  if (width - spaces[[maxPlace]] >= spaces[[minPlace]]) {
    spaces[[maxPlace]] <- spaces[[maxPlace]] + spaces[[minPlace]]
    spaces[[minPlace]] <- 0
  } else {
    spaces[[minPlace]] <-
      spaces[[minPlace]] + spaces[[maxPlace]] - width
    spaces[[maxPlace]] <- width
  }
  
}
spaces
```

Redo each list of how many squares for each color
```{r}
spareRow <- t(data.frame(rep(NA, width)))
rownames(spareRow) <- c()

if (spaces[[1]] != 0) {
  if (nrow(lastRed) == 0)
    lastRed <- rbind(lastRed, spareRow[])
  else
    lastRed[,] <- NA
  lastRed[, 1:spaces[[1]]] <- 1
}
if (spaces[[2]] != 0) {
  if (nrow(lastOrange) == 0)
    lastOrange <- rbind(lastOrange, spareRow[])
  else
    lastOrange[,] <- NA
  lastOrange[, 1:spaces[[2]]] <- 1
}
if (spaces[[3]] != 0) {
  if (nrow(lastYellow) == 0)
    lastYellow <- rbind(lastYellow, spareRow[])
  else
    lastYellow[,] <- NA
  lastYellow[, 1:spaces[[3]]] <- 1
}
if (spaces[[4]] != 0) {
  if (nrow(lastGreen) == 0)
    lastGreen <- rbind(lastGreen, spareRow[])
  else
    lastGreen[,] <- NA
  lastGreen[, 1:spaces[[4]]] <- 1
}
if (spaces[[5]] != 0) {
  if (nrow(lastBlue) == 0)
    lastBlue <- rbind(lastBlue, spareRow[])
  else
    lastBlue[,] <- NA
  lastBlue[, 1:spaces[[5]]] <- 1
}
if (spaces[[6]] != 0) {
  if (nrow(lastIndigo) == 0)
    lastIndigo <- rbind(lastIndigo, spareRow[])
  else
    lastIndigo[,] <- NA
  lastIndigo[, 1:spaces[[6]]] <- 1
}
if (spaces[[7]] != 0) {
  if (nrow(lastViolet) == 0)
    lastViolet <- rbind(lastViolet, spareRow[])
  else
    lastViolet[,] <- NA
  lastViolet[, 1:spaces[[7]]] <- 1
}
```

Replace the last row with the new one
```{r}
redDF[nrow(redDF),] <- lastRed
orangeDF[nrow(orangeDF),] <- lastOrange
yellowDF[nrow(yellowDF),] <- lastYellow
greenDF[nrow(greenDF),] <- lastGreen
blueDF[nrow(blueDF),] <- lastBlue
indigoDF[nrow(indigoDF),] <- lastIndigo
violetDF[nrow(violetDF),] <- lastViolet

colorList <- list(redDF, orangeDF, yellowDF, greenDF, blueDF, indigoDF, violetDF)
```


Regenerate the list of box sizes for each color
```{r, warning=FALSE}
redSize <- c()
orangeSize <- c()
yellowSize <- c()
greenSize <- c()
blueSize <- c()
indigoSize <- c()
violetSize <- c()
whiteSize <- c()

colorSize <-
  list(
    redSize,
    orangeSize,
    yellowSize,
    greenSize,
    blueSize,
    indigoSize,
    violetSize
  )

for (i in 1:7) {
  for (r in 1:(nrow(colorList[[i]]))) {
    for (c in 1:(ncol(colorList[[i]]))) {
      if(!is.na(colorList[[i]][r,c])) {
        colorSize[[i]] <- append(colorSize[[i]], colorList[[i]][r,c])
        colorList[[i]][r:(r+colorList[[i]][r,c] - 1),c:(c + colorList[[i]][r,c] - 1)] <- NA
      }
    }
  }
}
```


```{r}
rowCount <- nrow(df) - 2
colCount <- ncol(df) - 2

df <- data.frame(matrix(NA, nrow=(rowCount + 2), ncol=(colCount + 2)))

df[c(1,nrow(df)),] <- "x"
df[,c(1,ncol(df))] <- "x"

colors <- c("r", "o", "y", "g", "b", "i", "v")
squaresRemaining <- c(length(colorSize[[1]]), length(colorSize[[2]]), length(colorSize[[3]]), length(colorSize[[4]]), length(colorSize[[5]]),
                      length(colorSize[[6]]), length(colorSize[[7]]))

#maxSize <- floor(sqrt(ncol(df) - 1))
fullFlag <- FALSE

sameCounter <- 0
sameCheck <- squaresRemaining

rowFlip <- TRUE
colorNum <- 1

while(!fullFlag) {
  if(rowFlip) {
    rowSeq <- 2:(nrow(df)-1)
    colSeq <- 2:(ncol(df)-1)
    rowFlip <- FALSE
  }
  else {
    rowSeq <- (nrow(df)-1):2
    colSeq <- (ncol(df)-1):2
    rowFlip <- TRUE
  }
  for(r in rowSeq) {
    for(c in colSeq) {
      
      if(is.na(df[r,c])) {
      validFlag <- FALSE
      
      # colorNum <- floor(runif(1, min = 1, max = length(colors) + 1))
      # color <- colors[colorNum]
      # match(numberDF[r,c], colorSize[[colorNum]])
      
      colorNum <- floor(runif(1, min = 1, max = length(colors) + 1))
      while(TRUE) {
        colorNum <- floor(runif(1, min = 1, max = length(colors) + 1))
        color <- colors[colorNum]
        numSpot <- match(numberDF[r,c], colorSize[[colorNum]])
        boxSize <- colorSize[[colorNum]][numSpot]
        if(!is.na(match(numberDF[r,c], colorSize[[colorNum]])) & !is.na(colorSize[[colorNum]] & squaresRemaining[[colorNum]] != 0))
          break
      }
      colorSize[[colorNum]] <- colorSize[[colorNum]][-numSpot]
      
      # color <- colors[colorNum]
      # numSpot <- match(numberDF[r,c], colorSize[[colorNum]])
      # boxSize <- colorSize[[colorNum]][numSpot]
      # colorSize[[colorNum]] <- colorSize[[colorNum]][-numSpot]
      # 
      # if(colorNum == 7)
      #  colorNum <- 1
      # else
      #  colorNum <- colorNum + 1
        
      #leftovers <- squaresRemaining[[colorNum]] - 1
      
      #if((r + boxSize) <= (nrow(df) - 1) & (c + boxSize) <= (ncol(df) - 1)) {
      if((r + boxSize) <= (nrow(df) - 1) & (c + boxSize) <= (ncol(df) - 1) & squaresRemaining[[colorNum]] > 0) {
        
        temp <- as.data.frame(df[(r-1):(r+boxSize+1), (c-1):(c+boxSize+1)])
        temp[2:(nrow(temp)-1),2:(ncol(temp)-1)] <- NA
        temp[temp == "x"] <- NA
        if(nrow(temp) == 1)
          temp[1,1] <- NA
        else
          temp[c(1,nrow(temp)), c(1,ncol(temp))] <- NA
        
        if(!(FALSE %in% is.na(as.data.frame(df[r:(r+boxSize), c:(c+boxSize)]))) & !is.element(color, unlist(temp))) {
            validFlag <- TRUE
            squaresRemaining[[colorNum]] <- squaresRemaining[[colorNum]] - 1
            df[r:(r+boxSize), c:(c+boxSize)] <- color
        }
      }
        
      #if(!is.na(squaresRemaining[colorNum]) & squaresRemaining[[colorNum]] <= (.75 * rowCount * colCount)) {
        #colors <- colors[-colorNum]
        #squaresRemaining <- squaresRemaining[-colorNum]
      #}
      
      }
      
    }
  }
  
  if(!(TRUE %in% is.na(df)))
    fullFlag <- TRUE
  
  if(sameCheck == squaresRemaining)
    sameCounter <- sameCounter + 1
  else {
    sameCheck <- squaresRemaining
    sameCounter <- 0
  }

}

vec <- c(1:10)
vec
vec[-5]
```



